Assessing Bridge Condition in New York State Using Multiple Linear Regression

### 2024-12-03 paper/paper.qmd
Update abstract, introduction part. Upload a draft of appendix

---
title: "Assessing Bridge Condition in New York State Using Multiple Linear Regression"
subtitle: "Age, Municipality, and Ownership as Factors Influencing Bridge Condition Ratings "
author: 
  - Ariel Xing
thanks: "Code and data are available at: [https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis](https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true 
number-sections: true
bibliography: references.bib
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

# Create a histogram for the Condition variable
ggplot(bridge_data, aes(x = Condition)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(1, 7)) + # Ensures x-axis shows 1 to 7
  labs(
    x = "Bridge Condition Rating",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

# Visualization: Histogram with mode markers and density curve
ggplot(bridge_data, aes(x = AgeAtInspection)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) + # Histogram
  geom_vline(xintercept = seq(25, 150, by = 25), color = "red", linetype = "dashed", size = 0.8) + # Vertical lines for modes
  geom_density(aes(y = ..count..), color = "blue", size = 1) + # Density curve scaled to histogram
  labs(
    x = "Bridge Ages at Inspection (Years)",
    y = "Frequency"
  ) +
  theme_minimal()

@@

# Model

@@

# Results
The linear regression model predicting bridge condition ratings estimates how factors： the bridge age at inspection, the located municipality and the ownership influence the condition of bridges.The model result is shown in @tbl-model-result
```{r}
#| label: tbl-model-result
#| tbl-cap: "Summary Table of Bridge Condition Model"
#| echo: false
#| tbl-pos: "H"

@@

# Discussion

@@

# Additional data details

>>>>

---
title: "Assessing Bridge Condition in New York State Using Multiple Linear Regression"
subtitle: "Age, Municipality, and Ownership as Factors Influencing Bridge Condition Ratings "
author: 
  - Ariel Xing
thanks: "Code and data are available at: [https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis](https://github.com/Jiaqi-Xing/NYS_Bridge_Condition_Analysis)."
date: today
date-format: long
abstract: "This study examines the factors influencing bridge conditions in New York State, focusing on the roles of age, municipality location, and ownership, using condition ratings provided by NYSDOT as a measure of bridge quality. Applying a multiple linear regression model, the analysis reveals that for every additional year of age, a bridge's condition rating decreases by 0.02 points, reflecting the progressive impact of aging. On average, bridges located in towns and villages have condition ratings that are 0.17 and 0.19 points higher, respectively, than those in cities. Additionally, bridges owned by NYSDOT and other entities have condition ratings that are 0.09 and 0.16 points lower, respectively, than those owned by municipalities. These findings highlight the importance of targeted maintenance strategies that account for geographic and administrative factors to enhance infrastructure sustainability and safety.
"
format: pdf
toc: true 
number-sections: true
bibliography: references.bib
---

@@

Bridges are fundamental to New York State’s infrastructure, supporting transportation, economic activity, and community connections. They span diverse regions, from bustling urban centers to remote rural areas, accommodating varied traffic loads and environmental conditions. Maintaining these structures is essential to ensure public safety, economic efficiency, and reliable mobility. However, challenges such as aging infrastructure, resource constraints, and differing administrative responsibilities highlight the importance of understanding the factors that affect bridge conditions.

The estimand in this study is the effect of three key predictors: bridge age, municipality type, and ownership, on bridge condition ratings. Using a multiple linear regression model, we analyze condition ratings from 17,502 bridges in New York State to quantify how these factors influence bridge deterioration.

The results indicate that age is a major driver of bridge deterioration, with condition ratings decreasing by 0.02 points for each additional year. On average, bridges in towns and villages have better condition ratings than those in cities, with scores that are 0.17 and 0.19 points higher, respectively. Ownership also affects condition ratings: municipally owned bridges perform best, while NYSDOT-owned bridges score 0.09 points lower, and bridges owned by other entities score 0.16 points lower. These findings demonstrate the combined influence of geographic location and administrative responsibility on bridge conditions.

This research highlights the importance of targeted maintenance strategies that address the specific needs of aging bridges across different geographic and administrative contexts. By identifying the primary factors affecting bridge conditions, the study supports more effective resource distribution and prioritization in infrastructure management. These results contribute to the long-term sustainability and safety of the state’s transportation network.

The remainder of this paper is structured as follows. @sec-data provides an overview of the dataset and variables used in the analysis. @sec-model outlines the regression model, its assumptions, and limitations. @sec-result presents the results, while @sec-discussion discusses their implications and offers recommendations for future research. @sec-methodology introduces an idealized methodology for collecting bridge traffic flow data, and @sec-model-diagnostics details the diagnostics performed to validate the model.

@@

# Create a histogram for the Condition variable
ggplot(bridge_data, aes(x = Condition)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(1, 7)) + # Ensures x-axis shows 1 to 7
  labs(
    x = "Bridge Condition Rating",
    y = "Count"
  ) +
  theme_minimal()
```

@@

# Visualization: Histogram with mode markers and density curve
ggplot(bridge_data, aes(x = AgeAtInspection)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) + # Histogram
  geom_vline(xintercept = seq(25, 150, by = 25), color = "red", linetype = "dashed", size = 0.8) + # Vertical lines for modes
  geom_density(aes(y = ..count..), color = "blue", size = 1) + # Density curve scaled to histogram
  labs(
    x = "Bridge Ages at Inspection (Years)",
    y = "Count"
  ) +
  theme_minimal()

@@

# Model {#sec-model}

@@

# Results {#sec-result}
The linear regression model predicting bridge condition ratings estimates how factors： the bridge age at inspection, the located municipality and the ownership influence the condition of bridges.The model result is shown in @tbl-model-result
```{r}
#| label: tbl-model-result
#| tbl-cap: "Summary Table of Bridge Condition Model"
#| echo: false
#| tbl-pos: "H"

@@

# Discussion {#sec-discussion}

@@

# Appendix: Idealized Methodology for Collecting Bridge Traffic Flow Data

## Overview
The objective of this methodology is to design a robust system for collecting traffic flow data across New York State bridges. This data serves as a foundation for understanding the interaction between traffic intensity and bridge conditions. By employing rigorous sampling strategies and advanced collection techniques, this methodology ensures that the collected data is representative and actionable for policy and maintenance decisions.

## Target Population
The target population encompasses all operational bridges in New York State. These include bridges in urban, suburban, and rural areas, varying by traffic volume and ownership. Bridges were selected to account for:

- **Geographic Coverage**: Urban centers with high traffic density, suburban transit hubs, and rural connectors.
- **Administrative Ownership**: Municipal, NYSDOT, and privately owned bridges.
- **Traffic Intensity**: Low-traffic rural bridges, moderate suburban thoroughfares, and high-traffic urban crossings.

This definition ensures inclusivity, enabling the methodology to capture regional differences in traffic flow and their effect on bridge conditions.

## Sampling Approach

### Stratification Variables
To ensure the data represents the diversity of bridges, stratified random sampling is applied. Stratification variables include:

- **Municipality Type**: City, town, village.
- **Ownership**: Municipal, state, private.
- **Traffic Volume**: High, medium, low daily use.

Each bridge within a stratum is given an equal chance of selection, ensuring proportional representation across categories.

### Sample Size
Using the formula for sample size determination:

$$
n = \frac{Z^2 \cdot p \cdot (1-p)}{e^2}
$$

Where:

- \( Z = 1.96 \) for 95% confidence,
- \( p = 0.5 \), assuming maximum variability,
- \( e = 0.05 \), margin of error.

A minimum of 384 bridges is required to ensure statistically reliable data for the entire state.

## Data Collection Techniques

### Automatic Vehicle Counters (AVC)
- Deployed on selected bridges to continuously record vehicle counts and classifications.
- Cost-effective for high-traffic bridges requiring sustained monitoring.

### Camera-Based Systems
- Capture detailed traffic patterns, including speed and vehicle types, using AI-enhanced analysis.
- Effective for complex traffic conditions but requires privacy safeguards.

### GPS Data Integration
- Utilize anonymized GPS data to complement on-site measurements, especially for remote or less-monitored bridges.

### Manual Surveys
- Conducted periodically to validate automated data and address gaps, particularly in low-traffic or rural settings.

## Data Validation and Quality Control

### Attention Checks
Automated systems are periodically calibrated to ensure accuracy.

### Cross-Validation
Data from multiple sources (e.g., AVC, GPS) is compared to identify discrepancies.

### Consistency Checks
Traffic patterns are reviewed for anomalies that may indicate errors or data corruption.

## Challenges and Limitations

### Coverage Bias
Rural bridges may be underrepresented due to accessibility challenges. Tailored solutions are necessary to address this.

### Ethical Concerns
Privacy issues related to camera and GPS monitoring must be mitigated through anonymization protocols.

### Budget Constraints
High-cost technologies like AI-enabled cameras may limit widespread deployment.

## Relevance to Study
Incorporating traffic flow data enhances the predictive model of bridge conditions by linking usage intensity to structural deterioration. This data allows policymakers to prioritize maintenance based on a combination of bridge condition and traffic stress.
